Robust calibration of numerical models based on relative regret

### 2019-11-25 article.tex
remove part on weights of EM estimation, add k^*(u^i) with different u^i

{\color{red} kmean ou kexpectation ?}

@@

{\color{red} mean ou expectation ?}

@@

For the $J_{\mathrm{BHswap}}$ function, we can observe that those three methods give coherent results, as $\hatkmpe {}_{,\mathrm{KDE}} =\hatkmpe {}_{,\mathrm{EM}}=\hatkmpe {}_{,\mathrm{histogram}}\approx 0.8$. This is not the case for the  $J_{\mathrm{BH}}$ function however. Using Kernel density estimation (Gaussian), the estimation of $\kmpe$ is $\hatkmpe {}_{,\mathrm{KDE}} \approx 1.5$, while using the histogram and Gaussian mixture, $\hatkmpe {}_{,\mathrm{histogram}}=\hatkmpe {}_{,\mathrm{EM}}=3.8$.
This difference is explained by the accumulation of minimisers at this point: this challenges the assumption that $\mathbf{K}^*$ is continuous. As the density estimation techniques traditionally assume this continuity, the EM algorithm fits this using a normal distribution with a very small variance, while the KDE considers a sum of Gaussian kernels of constant bandwidth, located at the same point. This particular problem highlights the issue with $\kmpe$, as its estimation is highly sensitive to the density approximation procedure.

Alternatively, we can consider the mixing coefficients of the EM
density estimation. The Gaussian centered at $1.5$ has been attributed
a weight of $0.608$, while the one centered at $3.8$ has a mixing
weight of $0.226$. Ultimately, an optimiser would be more likely to be
close to $1.5$, but we do not have information on its performance in
this region around this value. \arthur{je ne comprends pas ce passage}

% as we can argue that both values may be a sensible choice for the MPE. This comes from the fact that the hypothesis that $\mathbf{K}^*$ is a continuous r.v.\ is violated. 
A way to avoid this issue is to introduce a bit of leeway for a value to be optimal, or at least ``acceptably not optimal''. This slackness takes the form of a relaxation coefficient, whose choice leads to a new family of robust estimators, along with a measure of their robustness via the coefficient.
%This leads us to the definition of a new family of estimators, for which one can attribute a {\it level of robustness} \elise{je me rend bien compte que cette phrase est un peu bizarre ...mais bon c'est l'idee}

@@

Similarly, if $p$ is fixed, the corresponding $\hat{\checka}_p$ is computed by searching for the smallest $\alpha$ such that $\max_{\mathbf{k}} \hat{\Gamma}_{\alpha}(\mathbf{k})$ is equal to the level $p$ fixed beforehand. In the end, if the value $\hat{\checka}_p$ is too large, the relaxation needed to get acceptable values with probability $p$ is very high, so the resulting estimation $\hat{\checkk}_p$ may not be relevant for the future application.
\victor{chapeau car ici je parle de l'estimation en pratique}

@@

The calibration of this numerical model is done in order to get a calibrated value $\kest$, that will be used for forecasts, using the same numerical model. We are now going to compare how well three of those values compare in a forecast context: $\kest_1$, $\hatkmean$ and $\kest_{\mathrm{global}}$, as the first two refer quite naturally to the notion of robustness, and the latter is an example of an estimate that does not take into account the uncertainties.
\clearpage
\subsection{Assessing the quality of the forecast of the calibrated model}
The calibration of the model has been performed using by integrating the model on a time-period $[0, T]$, called assimilation window. 
We now want to compare the quality of the different forecasts, from different calibrated bottom friction.
Those forecasts result from the integration of the numerical model between the time $T$ and a time $T_{\mathrm{pred}}$.

Given the probabilistic nature of the environmental conditions $\mathbf{U}$, the forecasts will also be probabilistic. We will then compare $\mathcal{M}_{\mathrm{pred}}^o(\mathbf{U})$ and $\mathcal{M}_{\mathrm{pred}}(\mathbf{k},\mathbf{U})$, for a calibrated $\mathbf{k}$.

\subsubsection{Squared forecast error}
Given two environmental conditions $\mathbf{u}$, used to run the computer simulation and $\mathbf{u}'$, used to generate the observations, the squared forecast error for the parameter $\mathbf{k}$ is
\begin{equation}
\label{eq:squared_forecast_error}
S_{\mathrm{pred}}(\mathbf{k},\mathbf{u},\mathbf{u}') = \left(\mathcal{M}_{\mathrm{pred}}(\mathbf{k},\mathbf{u}) - \mathcal{M}^{o}_{\mathrm{pred}}(\mathbf{u}')\right)^2
\end{equation}
Averaging over $\mathbf{u}$, $\mathbf{u}'$ defines the mean squared forecast error, defined on every point of the spatial domain, and at every time-steps. This can be done using a Monte-Carlo approximation: as $\mathbf{u}$ and $\mathbf{u}'$ are iid, let us assume that we have a set of samples $\{\mathbf{u}^i\}_{1\leq i \leq N_{\mathbf{u}}}$. 
\begin{equation}
  \label{eq:squared_forecast_error_averaged}
  S(\mathbf{k}) = \frac{1}{N_{\mathbf{u}}^2 }\sum_{i=1}^{N_{\mathbf{u}}}\sum_{j=1}^{N_{\mathbf{u}}} S_{\mathrm{pred}}(\mathbf{k},\mathbf{u}^i,\mathbf{u}^j) = \frac{1}{N_{\mathbf{u}}^2 }\sum_{i=1}^{N_{\mathbf{u}}}\sum_{j=1}^{N_{\mathbf{u}}}  \left(\mathcal{M}_{\mathrm{pred}}(\mathbf{k},\mathbf{u}^i) - \mathcal{M}^{o}_{\mathrm{pred}}(\mathbf{u}^j)\right)^2
\end{equation}
The mean squared forecast error averaged over the whole space and over all the time-steps gives an indication on the overall prediction quality of the prediction given this metric, and are represented on the right plot of Figure~\ref{fig:forecast_squared_error}. We can see that $\kest_{\mathrm{global}}$ performs slightly better than $\kest_1$, itself performing slightly better than $\hatkmean$.
Averaging $S(\mathbf{\kest}$ over only the time steps between $T$ and $T_{\mathrm{pred}}$, we have an indication on the quality of the forecast in the squared sense depending on the spatial position, as seen on the left plot of Figure~\ref{fig:forecast_squared_error}.
\begin{figure*}[!h]
\centering
\includegraphics[width=.9\textwidth]{Figures/forecast_error.pdf}
\caption{Quality of forecast, depending on the calibrated parameter. The left figures show the metric averaged over time, while the right figure show the mean forecast error, averaged over time and space}
\label{fig:forecast_squared_error}
\end{figure*}
We can see that the mean forecast error squared, on the right side, averaged over time and space is the smallest for $\kest_{\mathrm{global}}$, while $\kest_{1.0}$ performs slightly better $\hatkmean$.
\begin{figure}[!h]
  \centering
  \includegraphics[width=.9\textwidth]{Figures/forecast_CRPS}
  \caption{CRPS computed for different calibrated parameters}
\label{fig:forecast_crps}
\end{figure}

% \begin{figure*}[!h]
% \centering
% \includegraphics[width=10cm]{Figures/distrib_L2norm.pdf}
% \caption{Distribution of the forecast squared error, averaged over the different values of the uncertain parameter in the numerical simulations \victor{doublon avec fig.\ref{fig:forecast_squared_error} ?}}
% \label{fig:forecast_squared_error_distrib}
% \end{figure*}

\subsubsection{Continuous Ranked Probability Score}
Given the random variables $\mathcal{M}_{\mathrm{pred}}(\kest, \mathbf{U})$ and $\mathcal{M}_{\mathrm{pred}}^o(\mathbf{U})$, representing the probabilistic forecast and the probabilistic observations, we can define the cumulative distribution functions (CDF) $F_{\mathrm{pred}}(\cdot, \kest)$ and $F^o_{\mathrm{pred}}(\cdot)$
The Continuous ranked probability score (CRPS) measures the squared difference between the predicted CDF $F_{\mathrm{pred}}$ using a calibrated value, and the CDF of the observations $F_{\mathrm{pred}}^o$. 
\begin{equation}
\label{eq:def_crps}
\mathrm{CRPS}(\mathbf{k}) = \int_{\mathbb{R}} (F_{\mathrm{pred}}(x,\mathbf{k}) - F^o_{\mathrm{pred}}(x))^2 \,\mathrm{d}x
\end{equation}

@@

Finally, the adaptibility of this method has to be studied in a higher dimensional setting. High-dimensional density estimation is a challenge in itself, but is necessary in order to tackle more realistic applications.
%% The following commands are for the statements about the availability of data sets and/or software code corresponding to the manuscript.
%% It is strongly recommended to make use of these sections in case data sets and/or software code have been part of your research the article is based on.

% \sampleavailability{TEXT} %% use this section when having geoscientific samples available

% \appendix
% \section{}    %% Appendix A

>>>>

@@

@@

For the $J_{\mathrm{BHswap}}$ function, we can observe that those three methods give coherent results, as $\hatkmpe {}_{,\mathrm{KDE}} =\hatkmpe {}_{,\mathrm{EM}}=\hatkmpe {}_{,\mathrm{histogram}}\approx 0.8$. This is not the case for the  $J_{\mathrm{BH}}$ function however. Using Kernel density estimation (Gaussian), the estimation of $\kmpe$ is $\hatkmpe {}_{,\mathrm{KDE}} \approx 1.5$, while using the histogram and Gaussian mixture, $\hatkmpe {}_{,\mathrm{histogram}}=\hatkmpe {}_{,\mathrm{EM}}=3.8$.
This difference in the estimation of the mode is explained by the accumulation of minimisers at this point: this challenges the assumption that $\mathbf{K}^*$ is continuous. As the density estimation techniques traditionally assume this continuity, the EM algorithm fits this using a normal distribution with a very small variance, while the KDE considers a sum of Gaussian kernels of constant bandwidth, located at the same point. This particular problem highlights the issue with $\kmpe$, as its estimation is highly sensitive to the density approximation procedure.
% Alternatively, when considering the mixing coefficients of the EM estimation, the Gaussian centered at $1.5$ has the largest weight, of $0.608$. In other words, according to this estimated density, the optimal value $K^*$ has a probability of $0.608$ of being distributed normally around $1.5$.
% Alternatively, we can consider the mixing coefficients of the EM
% density estimation. The Gaussian centered at $1.5$ has been attributed
% a weight of $0.608$, while the one centered at $3.8$ has a mixing
% weight of $0.226$. Ultimately, an optimiser would be more likely to be
% close to $1.5$, but we do not have information on its performance in
% this region around this value. \arthur{je ne comprends pas ce passage}

% as we can argue that both values may be a sensible choice for the MPE. This comes from the fact that the hypothesis that $\mathbf{K}^*$ is a continuous r.v.\ is violated.

A way to avoid this issue is to introduce a bit of leeway for a value to be optimal, or at least ``acceptably not optimal''. This slackness takes the form of a relaxation coefficient, whose choice leads to a new family of robust estimators, along with a measure of their robustness via the coefficient.
%This leads us to the definition of a new family of estimators, for which one can attribute a {\it level of robustness} \elise{je me rend bien compte que cette phrase est un peu bizarre ...mais bon c'est l'idee}

@@

Similarly, if $p$ is fixed, the corresponding $\hat{\checka}_p$ is computed by searching for the smallest $\alpha$ such that $\max_{\mathbf{k}} \hat{\Gamma}_{\alpha}(\mathbf{k})$ is equal to the level $p$ fixed beforehand. In the end, if the value $\hat{\checka}_p$ is too large, the relaxation needed to get acceptable values with probability $p$ is very high, so the resulting estimation $\hat{\checkk}_p$ may not be relevant for the future application.
% \victor{chapeau car ici je parle de l'estimation en pratique}

@@

The calibration of this numerical model is done in order to get a calibrated value $\kest$, that will be used for forecasts, using the same numerical model. We are now going to compare how well three of those values compare in a forecast context: $\kest_1$, $\hatkmean$ and $\kest_{\mathrm{global}}$, as the first two refer quite naturally to the notion of robustness, and the latter is an example of an estimate that does not take into account the uncertainties.
\clearpage
\subsection{Assessing the quality of the forecast of the calibrated model}
The calibration of the model has been performed using by integrating the model on a time-period $[0, T]$, called assimilation window. 
We now want to compare the quality of the different forecasts, from different calibrated bottom friction.
Those forecasts result from the integration of the numerical model between the time $T$ and a time $T_{\mathrm{pred}}$.
Given the probabilistic nature of the environmental conditions $\mathbf{U}$, the forecasts will also be probabilistic. We will then compare $\mathcal{M}_{\mathrm{pred}}^o(\mathbf{U})$ and $\mathcal{M}_{\mathrm{pred}}(\mathbf{k},\mathbf{U})$, for a calibrated $\mathbf{k}$. The metric of comparison will be the squared forecast error, and the Continuous Ranked Probability Score (CRPS). The calibrated values that are compared are the minimiser of the expectation $\hatkmean$, the minimiser of the variance $\hatkvar$, the relative regret of level $p=1$: $\kest_1$ as described in the previous section, but also the global minimiser $\kest_{\mathrm{global}}$, and some conditional minimisers $\kest^*(\mathbf{u}^i)$ where the chosen environmental variables are $\{\mathbf{u}^i\}_{1\leq i \leq 4} = \{0.5, 0.55, 0.65, 0.7\}$. $\kest^*(0.6)$ has been omitted due to the fact that it result in a value very similar to the minimum of the expectation. These have been included in order to compare the different metrics obtained without using a robust approach. \todo{à revoir}

\subsubsection{Squared forecast error}
Given two environmental conditions $\mathbf{u}$, used to run the computer simulation and $\mathbf{u}'$, used to generate the observations, the squared forecast error for the parameter $\mathbf{k}$ is
\begin{equation}
\label{eq:squared_forecast_error}
S_{\mathrm{pred}}(\mathbf{k},\mathbf{u},\mathbf{u}') = \left(\mathcal{M}_{\mathrm{pred}}(\mathbf{k},\mathbf{u}) - \mathcal{M}^{o}_{\mathrm{pred}}(\mathbf{u}')\right)^2
\end{equation}
Averaging over $\mathbf{u}$, $\mathbf{u}'$ defines the mean squared forecast error, defined on every point of the spatial domain, and at every time-steps. This can be done using a Monte-Carlo approximation: as $\mathbf{u}$ and $\mathbf{u}'$ are iid, let us assume that we have a set of samples $\{\mathbf{u}^i\}_{1\leq i \leq N_{\mathbf{u}}}$. The squared forecast error can then be approximated using this finite set of samples:
\begin{equation}
  \label{eq:squared_forecast_error_averaged}
  S(\mathbf{k}) = \frac{1}{N_{\mathbf{u}}^2 }\sum_{i=1}^{N_{\mathbf{u}}}\sum_{j=1}^{N_{\mathbf{u}}} S_{\mathrm{pred}}(\mathbf{k},\mathbf{u}^i,\mathbf{u}^j) = \frac{1}{N_{\mathbf{u}}^2 }\sum_{i=1}^{N_{\mathbf{u}}}\sum_{j=1}^{N_{\mathbf{u}}}  \left(\mathcal{M}_{\mathrm{pred}}(\mathbf{k},\mathbf{u}^i) - \mathcal{M}^{o}_{\mathrm{pred}}(\mathbf{u}^j)\right)^2
\end{equation}
The mean squared forecast error averaged over the whole space and over all the time-steps gives an indication on the overall prediction quality of the prediction given this metric, and are represented on the right plot of Figure~\ref{fig:forecast_squared_error}. We can see that $\kest_{\mathrm{global}}$ performs slightly better than $\kest_1$, itself performing slightly better than $\hatkmean$.
Averaging $S(\mathbf{\kest})$ over the time steps between $T$ and $T_{\mathrm{pred}}$, we have an indication on the quality of the forecast in the squared sense depending on the spatial position, as seen on the left plot of Figure~\ref{fig:forecast_squared_error}.
\begin{figure*}[!h]
\centering
\includegraphics[width=.9\textwidth]{Figures/L2_full.pdf}
\caption{Quality of forecast, depending on the calibrated parameter. The left figures show the metric averaged over time, while the right figure show the mean forecast error, averaged over time and space}
\label{fig:forecast_squared_error}
\end{figure*}
We can see that the mean forecast error squared, on the right side, averaged over time and space is the smallest for $\kest_{\mathrm{global}}$, while $\kest_{1.0}$ performs slightly better $\hatkmean$.

% \begin{figure*}[!h]
% \centering
% \includegraphics[width=10cm]{Figures/distrib_L2norm.pdf}
% \caption{Distribution of the forecast squared error, averaged over the different values of the uncertain parameter in the numerical simulations \victor{doublon avec fig.\ref{fig:forecast_squared_error} ?}}
% \label{fig:forecast_squared_error_distrib}
% \end{figure*}
One issue with the squared forecast error is that it will penalize strongly a high sensitivity to the environmental conditions \todo{à conserver/changer ?}. A way to deal with this is to use another metric, that takes the probabilistic nature of the forecasts into account.
\subsubsection{Continuous Ranked Probability Score}
Given the random variables $\mathcal{M}_{\mathrm{pred}}(\kest, \mathbf{U})$ and $\mathcal{M}_{\mathrm{pred}}^o(\mathbf{U})$, representing the probabilistic forecast and the probabilistic observations, we can define the cumulative distribution functions (CDF) $F_{\mathrm{pred}}(\cdot, \kest)$ and $F^o_{\mathrm{pred}}(\cdot)$
The Continuous ranked probability score (CRPS) measures the squared difference between the predicted CDF $F_{\mathrm{pred}}$ using a calibrated value, and the CDF of the observations $F_{\mathrm{pred}}^o$. 
\begin{equation}
\label{eq:def_crps}
\mathrm{CRPS}(\mathbf{k}) = \int_{\mathbb{R}} (F_{\mathrm{pred}}(x,\mathbf{k}) - F^o_{\mathrm{pred}}(x))^2 \,\mathrm{d}x
\end{equation}

\begin{figure}[!h]
  \centering
  \includegraphics[width=.9\textwidth]{Figures/CRPS_full}
  \caption{CRPS computed for different calibrated parameters}
\label{fig:forecast_crps}
\end{figure}

@@

Finally, the adaptibility of this method has to be studied in a higher dimensional setting. High-dimensional density estimation is a challenge in itself, but is necessary in order to tackle more realistic applications.
% \appendix
% \section{}    %% Appendix A
